---
title: "Central Tend Questions"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
knitr::opts_chunk$set(echo = FALSE)
lion <- read.csv("https://raw.githubusercontent.com/Gharv/Questions/master/lionCrater.csv")
rhino <- read.csv("https://raw.githubusercontent.com/Gharv/Questions/master/blackRhinoCrater.csv")
```


## Max & Min

*The lionCrater.csv and blackRhinoCrater.csv files have already been imported and saved with the names **lion** and **rhino** respectively*

Use the interactive code block below to answer the following questions:

```{r max, exercise=TRUE}

```

### Questions


```{r max, quiz}
quiz(
  question("What is the minimum and maximum population size for lions?",
    answer("1963 & 1992"),
    answer("72.50 & 69.20"),
    answer("12.00 & 123.00", correct = TRUE),
    answer("52.25 & 88.75")
  ),
  question("What is the minimum and maximum population size for black rhinos?",
    answer("1980 & 1995"),
    answer("11.00 & 16.00"),
    answer("10.00 & 21.00", correct = TRUE),
    answer("14.00 & 14.12")
  )
)
```


## Mean, Median & Standard Deviation


*The lionCrater.csv and blackRhinoCrater.csv files have already been imported and saved with the names **lion** and **rhino** respectively*

Use the interactive code block below to answer the following questions:

```{r mean, exercise=TRUE}

```

### Questions

```{r mean, quiz}
quiz(
  question("Which ones are the mean, median, and standard deviation of the population size of lions?",
    answer("69.20", correct = TRUE),
    answer("14.00"),
    answer("123.00"),
    answer("72.50", correct = TRUE),
    answer("27.81", correct = TRUE)
  ),
  question("Check the mean and median of the population size for black rhinos.",
    answer("1988"),
    answer("21.00"),
    answer("14.00", correct = TRUE),
    answer("69.20"),
    answer("14.12", correct = TRUE)
  )
)
```


## Quantiles 

*The lionCrater.csv and blackRhinoCrater.csv files have already been imported and saved with the names **lion** and **rhino** respectively*

Use the interactive code block below to answer the following questions:

```{r quantile, exercise=TRUE}

```

### Questions

```{r quantile, quiz}
quiz(
  question("Which are the quantile (0%,25%,50%,75%,100%) population sizes for black rhinos?",
    answer("10.00", correct = TRUE),
    answer("11.00", correct = TRUE),
    answer("14.00", correct = TRUE),
    answer("14.12"),
    answer("16.00", correct = TRUE),
    answer("21.00", correct = TRUE)
  )
)
```

