---
title: "Central Tendency (R Instructional Worksheet)"
output:
  tufte::tufte_handout:
    toc: yes
    toc_depth: 1
  tufte::tufte_html:
    toc: yes
    toc_depth: 1
---
NOTES: DONT HAVE CHIMP DATA


#Descriptive Statistics

We will use the chimpanzee dataset that we imported earlier to explore basic descriptive statistics. But first, we need to look at the structure of the dataset.  

```{r eval=FALSE}
str(chimp)
```

This dataset has two variables, Year and Population. Let’s explore the Population variable.  

```{r eval=FALSE}
chimp$Population #display the data in the Population column
```

Minimum – smallest value  
```{r eval=FALSE}
min(chimp$Population)
```

Maximum – largest value
```{r eval=FALSE}
max(chimp$Population)
```

Range (Minimum and Maximum)
```{r eval=FALSE}
range(chimp$Population)
```

Median – middle number when all of the values are organized from smallest to largest
```{r eval=FALSE}
median(chimp$Population)
```

Quantile
```{r eval=FALSE}
quantile(chimp$Population)
```

Mean is a measure of central tendency of the data, and the code to calculate it in R is as follows:  
```{r eval=FALSE}
mean(chimp$Population)
```

The summary command calculates a group of the descriptive statistics all at once.
```{r eval=FALSE}
summary(chimp$Population)
```

All of the values are rounded to the nearest integer before these values are calculated, which is why they are slightly different than the values calculated above.


#Missing Values

Datasets are rarely perfect, and there is often data missing. First, here is a simple example with one value missing.

```{r}
a <- c(1,2,3,NA,5)
mean(a)
mean(a, na.rm = TRUE)
```

Next, let’s us see how to deal with missing data in a real dataset. Let us import the ‘lion’ dataset. This dataset has the population number of lions in the Serengeti Plains from 1966-1990.  
First import the dataset and name it ‘lion’.

```{r}
lion <- read.csv("lion.csv", header = T)
```

If you look at the data, you’ll see that there are no population sizes for the years 1969-1973.  
Now, what happens if we try to find the mean of the population size?

```{r}
mean(lion$Population)
```

We can use the trick we just learned to get the mean population size.  

```{r}
mean(lion$Population, na.rm = TRUE)
```

However, instead of having to include the ’na.rm’ argument every time, we can just delete the rows that are missing data.

```{r}
lion2 <- na.omit(lion)
```

\newpage

#Problems

1. Import the Lion and Black Rhino files into R that were used in previous assignments (‘lionCrater.csv’ and ‘blackRhinoCrater.csv’).  
\vspace*{1\baselineskip}
2. What is the minimum and maximum population size for lions?  
\vspace*{1\baselineskip}
3. What is the mean, median, and standard deviation of the population size of lions?  
\vspace*{1\baselineskip}
4. What is the minimum and maximum population size for black rhinos?  
\vspace*{1\baselineskip}
5. What are the quantile (0%,25%,50%,75%,100%) population sizes for black rhinos?  
\vspace*{1\baselineskip}
6. Use the *summary* command to find the descriptive statistics of population size for black
rhinos. What is the mean and median?  
















