---
title: "Correlation"
output:
  tufte::tufte_handout:
    toc: yes
    toc_depth: 1
  tufte::tufte_html:
    toc: yes
    toc_depth: 1
---

#Correlation

We will use the iris data to look at the relationship between petal length and width. (The iris dataset is part of the base R installation and can be referenced directly, as shown in the sample code for this section.)  

First, we can look at the relationship on a graph.  

```{r fig.fullwidth=TRUE}
par(mar=(c(2,2,2,2)))
plot(iris$Petal.Width, iris$Petal.Length, main = "Petal Width vs. Length", xlab = "Petal Width", ylab = "Petal Length")
```

The plot shows that there is a positive, linear relationship between petal length and width.  

If we want to know if the relationship is significant, we can calculate the correlation between the two variables.  

```{r}
cor.test(iris$Petal.Width, iris$Petal.Length)
```

Since the p-value is less than 0.05 there is a significant correlation between these two variables. The variables are 96% correlated.



