---
title: "Linear Regression"
output:
  tufte::tufte_handout:
    toc: yes
    toc_depth: 1
  tufte::tufte_html:
    toc: yes
    toc_depth: 1
---

#Linear Regression

The goal of Linear Regression is to model a linear relationship between an independent (X) and a dependent (Y) variable. If we know the petal width can we estimate the petal length? (The iris dataset is installed as part of the base R installation and can therefore be referenced directly, as in the sample code for this section.)  
\vspace*{1\baselineskip}
First, fit a linear model.  

```{r}
model <- lm(iris$Petal.Length ~ iris$Petal.Width)
```

Summarize the model.

```{r}
summary(model)
```

So, our predictive equation is:  
Petal Length = 1.08356 + 2.22994(Petal Width)  
\vspace*{1\baselineskip}
These results tell us that:  

* the intercept is significantly different than 0 (P<0.05)
* the slope is significantly different than 0 (P <0.05)
* the model is a good fit for the data (R 2 = 0.9271). This means that 93% of the variation in petal length is explained by the petal width.  

We can also add a linear regression line directly in the plot of the two variables:

```{r  fig.width = 4, fig.height = 3, fig.fullwidth = TRUE}
#par(mar=(c(5,4,4,2)))
plot(iris$Petal.Width, iris$Petal.Length, main = "Petal Width vs. Length", xlab =
"Petal Width", ylab = "Petal Length")
abline(model)
```









