---
title: "Variance"
output:
  tufte::tufte_handout:
    toc: yes
    toc_depth: 1
  tufte::tufte_html:
    toc: yes
    toc_depth: 1
---

#Measuring Variability

We will use the chimpanzee dataset that was introduced in section 2.6. Please review that section if you need to refresh your understanding of how to import this dataset into R.  

Variance is a measure of spread of the data, and is calculated in R as follows:  

```{r eval=FALSE}
var(chimp$Population)
```

Standard deviation is just the square root of variance (how much the values typically vary from the average value), and is calculated in R as follows:

```{r eval=FALSE}
sd(chimp$Population)
```

#Outliers and Variance


\newpage

#Problems

1. Import the Lion and Black Rhino files into R that we were using in the last assignment (‘lionCrater.csv’ and ‘blackRhinoCrater.csv’).  
\vspace*{1\baselineskip}
2. What is the variance and standard deviation of population size for black rhinos?  

